---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## tabulergm

<!-- badges: start -->
[![R-CMD-check](https://github.com/gvegayon/tabulergm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gvegayon/tabulergm/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`tabulergm` is an R package designed to translate models estimated with the
[`ergm`](https://cran.r-project.org/package=ergm) framework into
publication-ready tables and explanatory summaries.

## Installation

```r
# Install from GitHub
# install.packages("remotes")
remotes::install_github("gvegayon/tabulergm")
```

## Example

Here is a simple example fitting an ERGM to the Florentine marriage network and
generating a summary table:

```{r}
#| label: example
#| message: false
#| warning: false
library(ergm)
library(tabulergm)

# Fit a simple ERGM
data(florentine)
model <- ergm(
  flomarriage ~ edges + triangle,
  control = control.ergm(seed = 42)
)

# Parse the model
model_terms <- parse_ergm_model(model)
model_terms[, c("term", "estimate", "se", "pvalue", "description")]
```

The term dictionary also includes mode-specific terms for bipartite ERGMs.  A
formula is enough to inspect the available metadata before fitting a model:

```{r}
#| label: bipartite-terms
#| message: false
#| warning: false
bipartite_terms <- parse_ergm_formula(
  network ~
    gwb1dsp(0.5, fixed = TRUE) + gwb2dsp(0.5, fixed = TRUE) +
    b1factor("type") + b2factor("group") +
    b1nodematch("type") + b2nodematch("group")
)

bipartite_terms[, c("term", "attribute", "description")]
```

## License

MIT © George Vega Yon
