---
title: "Fitting an ERGM and Viewing the Output with tabulergm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting an ERGM and Viewing the Output with tabulergm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>"
)
```

## Overview

This vignette demonstrates how to:

1. Fit an ERGM using the **ergm** package.
2. Produce a publication-ready Markdown table with `tabulergm_table(format = "markdown")`.
3. Render the table inline in a Quarto or R Markdown document using `results: asis`.
4. Interactively preview the table in the RStudio viewer (or a browser) with
   `tabulergm_view()`.

## Setup

```{r load-packages, message = FALSE, warning = FALSE}
library(ergm)
library(tabulergm)
```

## Fitting the model

We use the Florentine marriage network, which ships with **ergm**, and fit a
simple model with an `edges` term and a `nodematch` term for wealth quartile.

```{r fit-model, message = FALSE, warning = FALSE}
data(florentine)

model <- ergm(
  flomarriage ~ edges + nodematch("wealth"),
  control = control.ergm(seed = 42)
)
summary(model)
```

## Creating a Markdown table

Calling `tabulergm_table()` with `format = "markdown"` returns a
`knitr_kable` object.  Adding the chunk option `results: asis` (or
`results = "asis"` in R Markdown) causes knitr to emit the table verbatim,
so the Markdown renderer (Quarto, Pandoc, GitHub, etc.) formats it properly.

Math notation in the **math** column is automatically wrapped in `$$...$$` so
that MathJax or Pandoc can render it as display LaTeX.  Network figures in the
**figure** column are converted to `<img>` HTML tags.

> **Quarto tip:** use `#| results: asis` (or `results = "asis"` in R Markdown)
> on the chunk so that knitr emits the table verbatim instead of quoting it.

```{r markdown-table, results = "asis"}
tabulergm_table(
  model,
  include_math        = TRUE,
  include_description = TRUE,
  format              = "markdown"
)
```

## Inspecting a formula without a fitted model

You can also pass a bare formula to inspect term metadata before fitting:

```{r formula-table, results = "asis"}
tabulergm_table(
  flomarriage ~ edges + nodematch("wealth") + triangle,
  format = "markdown"
)
```

## Interactive preview with `tabulergm_view()`

During an interactive session you can call `tabulergm_view()` to open the
table in the RStudio viewer pane or the system browser:

```{r view-demo, eval = FALSE}
tabulergm_view(model, include_math = TRUE, include_description = TRUE)
```

`tabulergm_view()` builds a self-contained HTML page that loads MathJax from
a CDN, so LaTeX math and embedded network figures render immediately without
any additional setup.
